\chapter{ Validation } \label{chap:validation}


Validating the collected data and the subsequently identified patterns cannot be made in a formal way.

The only course of action available is therefore to measure the effects that applying the gathered information has in a real world situation. In order to do so, measurements must be made before and after the application.

\section{Criteria}

Having a wide are of application and focusing its attention in several areas and domains, measuring the success of adopting DevOps can be difficult. In this work the evaluation of the success factor will be divided into two main categories having each a set of criteria for acceptance:


	\begin{itemize}
        \item{ Technical aspects }
        	\begin{itemize}
				\item{Deployment Frequency} - Deploying an application is the process by which software can be made available to the customer. By increasing the deployment frequency teams can more frequently transmit value to the customer. 
                
				\item{Deployment Time} - Deployment time is of great importance as it enables companies to quickly respond to change. 
                
				\item{Infrastructure Elasticity} - Being able to quickly and in an automated way increase or decrease the infrastructure size can mean the difference between success and failure. 
                
			\end{itemize}
            
            
        \item{ Organizational aspects }
        	\begin{itemize}
				\item{Team satisfaction} - Team satisfaction is essential as it promotes more motivated professionals and better results.
                
				\item{ Managerial satisfaction } - The satisfaction of the managerial team is of great importance as it usually reflects the current state of the business. 
                
			\end{itemize}

	\end{itemize}


	For the previous criteria the following acceptance and rejection metrics will be used: 
    
		\begin{table}[h!]
			\centering
            \caption{Technological Criteria}
			\label{tab1}
			\begin{tabular}{|l|c|c|}
               	\hline
 				Criteria & Acceptance & Rejection  \\ \hline
 				Deployment Frequency & Increase  & Decrease   \\ \hline
 				Deployment Time & Decrease &  Increase  \\ \hline
 				Infrastructure Elasticity & Decrease & Increase \\ \hline 
			\end{tabular}
		\end{table}



		\begin{table}[h!]
			\centering
            \caption{Organizational Criteria}
			\label{tab2}
				\begin{tabular}{|l|c|c|}
                	\hline
 					Criteria & Acceptance & Rejection  \\ \hline
 					Team satisfaction & Increase  & Decrease   \\ \hline
 					Managerial Satisfaction & Increase &  Decrease  \\ \hline
 					Infrastructure Elasticity ( time to provision new resources) & Decrease & Increase \\ \hline 
				\end{tabular}
			\end{table}

	